### 2018-03-13 tex/merf/TAL/conclusion.tex
adding merf

In this work, we present a morphology-based entity and relational entity extraction framework for Arabic text.
\framework provides a friendly interface where the user defines tag types 
and associates them with regular expressions defined over Boolean formulae.
The Boolean formulae are in turn defined over matches of Arabic morphological features and
a novel extended synonymy feature ($Syn^k$).
\framework allows the user to associate code actions with each regular sub-expression 
and to define semantic relations between sub-expressions. 
%\framework uses Sarf, an Arabic morphological analyzer, 
%to compute morphological and thereafter
%regular expression matches, and relational entities. 
We evaluate \framework with several case studies and compare with existing application-specific 
techniques.
The results show that \framework requires shorter development time and effort compared 
to existing techniques and produces reasonably accurate results within a reasonable 
overhead in run time. 
In the future, \framework will support user-defined cross-reference predicates, 
and will infer morphological features from relevant example words to express a concept.

%Currently, \framework supports one built in cross-reference predicate based on the $Syn^2$ feature. 
%In the future, \framework will support user-defined cross-reference predicates. 
%Currently the user selects the morphological features to specify the MBF. 
%We will explore techniques that can infer the features from example words that the user judges as relevant to the basic concept in question.