### 2018-03-13 tex/merf/TALMajor/answer/Reviewer1.comments.tex

\section*{Reviewer 1 comments}
\textit{Comments to the Author}\\

The paper describes a tool (MERF) for entity and relation 
extraction in Arabic text.
The tool allows to describe regular expressions which atomic 
elements are morphological descriptions of Arabic words.
The regular expressions are compiled as finite state 
automata that are used for pattern matching.

MERF uses a morphological analyzer for Arabic Sarf that 
is itself based on finite state transducers.
MERF proposes several GUI to edit rules and analyze results.
The paper is clearly written.

\answer{
  Thank you. 
}

\begin{enumerate}[leftmargin=0mm,label=\bfseries CommentR1.\arabic*]

  \item \label{Review1.1} 
The authors chose to use formal notation to describe the 
objects manipulated by the software. 
This make things rigouros but the notations are sometimes heavy.
\answer{
  We clarified the following notations.... 
  The new notations appear in Sectioon AA in teh paper. 
}

\item \label{Review1.2} 
I did not understand what is the exact status of the predicate 
isA and how it is used, the authors should be a bit 
 clearer about that.

\answer{
We modified the description of isA to .... 
}

\item \label{Review1.3} 
The system is evaluated and compared to four other systems for 
tasks of Information Extraction.
the data used for the evaluation is not clearly described, 
did you have a manual gold annotation of all the data ?
\answer{
We have to look at the test data for the paper. 
}

\item \label{Review1.4} 
what is the size of the evaluation data ? 
did you have a train and a test data set ?
\answer{
No training... testing yes. 
}

\item \label{Review1.5} 
how did you compute the time for the implementation of rules 
with the other system (one of the tasks needed 2 month of work 
!! is it really 2 acutal months of work or is it the period of 
time during which the development has been done)
\answer{
A research assistant 6 hours course work and 14 hours 
teaching duties performed the task in two months. 
}

\item \label{Review1.5} 
more should be said about the evaluation methodology.
\answer{
We added a methodology section. 
}

\item \label{Review1.6}
you say nothing about the availability of the system, 
is it distributed ? free ? what kind of licence ?
\answer{
Available online on github, with GPL license. 
}

\end{enumerate}