### 2018-03-13 tex/merf/TALMajor/answer/Reviewer4.comments.tex

\section*{Reviewer 4 comments}
\textit{Comments to the Author}\\

This work presents MERF; a GUI-enabled regular expressions 
based tool for entity and relational entity extraction, 
enhanced for Arabic text by enabling morphology-based matching.

\begin{enumerate}[leftmargin=0mm,label=\bfseries CommentR2.\arabic*]

\item \label{Review.4.1} 
Clarity:   
It is not clear how to associate code actions with 
subexpressions!
\answer{
  We clarify that... 
}

\item \label{Review.4.1} 
Correctness :  
- $Syn^k$ introduces a lot of noise especially if $k$ is large. 
\answer{
  7 is a limit... 
}

\item \label{Review.4.1} 
In the results section, there is no standard evaluation. 
The evaluation is based on comparisons with non-standard 
task-specific tools. 
Also, the reported precision and recall are very sensitive 
to the quality of the annotator!
\answer{
  Same answer to reviewer 3 results... 
}

\end{enumerate}