### 2018-03-13 tex/merf/TALMajor/answer/additionalComments.tex

\section*{Additional Comments:} 
Please consider the following related efforts:

Y Benajiba et al. 2007. Anersys: An arabic named entity recognition system based on maximum entropy

Y Benajiba et al. 2008. Arabic named entity recognition using optimized feature sets.

A Pasha, et al. 2014. MADAMIRA: A Fast, Comprehensive Tool for Morphological Analysis and Disambiguation of Arabic.

A Shahrour et al. 2016. CamelParser: A System for Arabic Syntactic Analysis and Morphological Disambiguation.

\answer{
We discuss the work in the related section. 
}