### 2018-03-13 tex/merf/TALMajor/answer/answer.tex

\documentclass[11pt]{article}
\usepackage[letterpaper,margin=1.5in,left=35mm,right=20mm, top=20mm, bottom=20mm ]{geometry}
\input{packages}

\tcbuselibrary{skins}
\usetikzlibrary{shadings}
\input{answerMacros}
\input{answerPackages}

\title{MERF: Morphology-based Entity and Relational Entity Extraction Framework for Arabic
 \\
~~ \\
  Response to Major Revision }
\date{}

\begin{document}
\maketitle

{\noindent Dear Editor-in-Chief, Associate Editor and expert Reviewers,}

We would like to thank you for your recommendations, valuable comments, 
and constructive observations. 
%
In this revised version of the paper we made every effort to address the 
concerns brought up in the reviews. 
%
We highlighted the changed parts in the paper with color blue. 
This document discusses the comments of the Reviewers 
in details, provides answers for them, and points to sections of the revised 
paper where the comments were addressed. 
\\
\noindent
Thank you.
\newline
\newline
\hrule
\newpage

\input{Reviewer1.comments}
~\newline
\hrule

\input{Reviewer2.comments}
~\newline
\hrule

\input{Reviewer3.comments}
~\newline
\hrule

\input{Reviewer4.comments}
~\newline
\hrule

\input{additionalComments}

%\input{other_notes}

%\newsavebox\mytempbib
%\savebox\mytempbib{\parbox{\textwidth}{
%\input{psqBMCaig.bbl}
%}}

\end{document}