### 2018-03-13 tex/merf/TALMajor/answer/editorComments.tex

\section*{Editor-in-Chief Letter}

Dear Amin Jaber,

We are happy to inform you that your submission to the journal 
``Traitement automatique de l'arabe et des langues apparentées (TAL 58-3)''
has been accepted but requires major revisions.

You are invited to make major revisions to be considered for publication in 
the second review round.

Please submit your revised article before April 2, 2018 (midnight Paris time).

The reference of your submission is:
MERF: Morphology-based Entity and Relational Entity Extraction Framework for Arabic, 
A. Jaber [et al.] (sciencesconf.org:tal-58-3:171885)

Best Regards,
tal-58-3 team

https://tal-58-3.sciencesconf.org