### 2018-03-13 tex/merf/TALMajor/macros.tex

\documentclass[english,utf8x]{article-hermes}

%%%%%%%% packages %%%%%%%%%%
\usepackage{xspace}
\usepackage{arabtex}
\usepackage{array}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0},
}
\usepackage{booktabs}
\usepackage[ruled]{algorithm2e}
\usepackage{etex}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{graphics}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{savesym}
\savesymbol{bigtimes}
\usepackage{mathabx}
\restoresymbol{TXF}{bigtimes}
\usepackage{utf8}
\usepackage{subfigure}
\usepackage{todonotes}
\usetikzlibrary{shapes,arrows}
\usepackage{color,soul}
\usepackage{verbatim}
\usepackage{array}
\usepackage{multicol}
\usepackage{fancyvrb}
\usepackage{relsize}
\usepackage{wasysym}
\usepackage{pbox}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%% Macros %%%%%%%%%%
\def\framework{\textsc{MERF}\xspace}
\newcommand{\itodo}[1]{\todo[inline,color=green]{\tiny #1}}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcommand{\TinyRL}[1]{{\RL{#1}}}
\newcommand{\cci}[1]{{\small \texttt{#1}}}
%%% RL commands
\newcommand{\utfrl}[1]{\setcode{utf8}\RL{#1}\setcode{standard}}
\newcommand{\notrutfrl}[1]{\transfalse\setcode{utf8}\RL{#1}\setcode{standard}\transtrue}
\newcommand{\notrrl}[1]{\transfalse\RL{#1}\transtrue}
\newcommand{\noarrl}[1]{\arabfalse\RL{#1}\arabtrue}
\def\showvrb#1{%
\texttt{\detokenize{#1}}%
}
%%% define cal
\makeatletter
\DeclareRobustCommand*\cal{\@fontswitch\relax\mathcal}
\makeatother
% done
%other 
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\submitted[15/10/2017]{TAL~58-03}{R}

\title[\framework]{\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic}

%\author{Firstname$_1$ Lastname$_1$\fup{*} \andauthor Firstname$_2$ Lastname$_2$\fup{**}  } 
%\address{%
%\fup{*} First affiliation\\
%\fup{**} Second affiliation
%}

\author{}
%\author{Amin A. Jaber \andauthor Fadi A. Zaraket}

\address{}
%\address{American University of Beirut, Beirut 1107 2020, Lebanon}

\abstract{
\input{abstract}
}

\keywords{
Arabic,
information extraction,
natural language processing,
entity,
relational entity,
information visualization,
tagging.
}

\motscles{
%mot-cl\'e$_1$,
%mot-cl\'e$_2$,
%mot-cl\'e$_3$.
}

\resume{
%Le r\'esum\'e est \`a \'ecrire ici.
}

\begin{document}

\maketitlepage