### 2018-03-13 tex/merf/merf/merf.tex

\include{macros}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic}
%~\IEEEmembership{Member,~IEEE,}
\author{Ameen~Jaber,
        Fadi~A.~Zaraket
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem A. Jaber and F. Zaraket are with the Department
of Electrical and Computer Engineering, American University of Beirut, Beirut,
Lebanon.\protect\\
E-mail: \{aaj15,fz11\}@aub.edu.lb}}

\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
\input{abstract}
\end{abstract}

\begin{IEEEkeywords}
Tagging, Arabic, Natural language processing, Morphological analysis, Computational linguistics, Information extraction, Entity, Relational Entity.
\end{IEEEkeywords}}

% make the title area
\maketitle

% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}
\input{introduction}
%\hfill mds
%\hfill August 15, 2013
\section{Motivation}
\label{sec:motivation}
\input{motivation}

\section{Overview}
\label{sec:overview}
\input{overview}

\section{Preliminaries}
\label{sec:preliminaries}
\input{preliminaries}

\section{Related Work}
\label{sec:related}
\input{relatedwork}

\section{\framework}
\label{sec:framework}
\input{tool}

\section{\framework GUI}
\label{sec:gui}
\input{gui}

\section{Implementation}
\label{sec:implementation}
\input{implementation}

\section{Results}
\label{sec:results}
\input{results}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

\bibliography{references}
\end{document}