### 2018-03-13 tex/merf/merf/overview.tex

%\begin{figure}[h!]
%  \centering
%  \includegraphics[width=0.45\textwidth]{figures/overview}
%  \caption{\framework Overview}
%  \label{f:overview}
%\end{figure}

\begin{figure}[h!]
\centering
\resizebox{\columnwidth}{!}
{\tiny \input{figures/overview.tex}}
\caption{\framework flow diagram.}
\label{f:overview}
\end{figure}

Figure~\ref{f:overview} shows the flow diagram of \framework.
\framework takes as input the Arabic text. 
The user interacts with \framework through a user-friendly interface to specify the
Morphology-based Boolean formulae, the Morphology-based regular expressions, 
the relation declarations, and the actions.
\framework analyzes the text using an inhouse Arabic morphological anaylzer~\cite{sarf} and computes morphological solutions
for each word in the Arabic text.
\framework passes the solutions to the Boolean formula simulator including the K-Synonym detector 
which computes the matches and produces a tag set for each word.
\framework then passes the sequence of tag sets to the regular expression simulator which computes 
chunks; i.e. sequences of words, whose tag sets match the expressions. 
\framework then evaluates the semantic relation declarations against the matches to compute the semantic relational entities, 
and executes the corresponding user-defined actions. 
Finally, \framework uses a cross-referrence relation such as the \cci{isA} relation to create relations across chunks and 
documents.

\framework presents the resulting tags to the user incrementally in the form of 
text annotated with style and color legends, match trees, and graphs.
\framework also provides statistical analysis tools that help compare sets of tags to reference tags,
compute standard accuracy measures, and correct generated annotations.