### 2018-03-14 tex/merf/OldSubmission/CoverLetter/coverletter.tex

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm}

\usepackage{charter}

\newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
\includegraphics[width=1.2\linewidth]{aub-logo.jpg}}} % Company/institution logo at the top left of the page
\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=50pt} % Slightly decrease the height of the signature field
\newlfmP{addrfromphone} % Print a phone number under the sender's address
\newlfmP{addrfromemail} % Print an email address under the sender's address
\PhrPhone{Phone} % Customize the "Telephone" text
\PhrEmail{Email} % Customize the "E-mail" text

\lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{Ameen Jaber} % Name

\addrfrom{
\today\\[12pt] % Date
American University of Beirut\\
P.O.Box 11-3875\\
Riad El-Solh / Beirut 1107 2020
}

\phonefrom{(+961)70-962096} % Phone number

\emailfrom{aaj15@aub.edu.lb} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear Jian Pei, editor in cheif, }% Greeting text
\closeline{Sincerely yours,} % Closing text

%\nameto{Mrs. Jane Smith} % Addressee of the letter above the to address

%\addrto{
%Recruitment Officer \\ % To address
%The Corporation \\
%123 Pleasant Lane \\
%City, State 12345
%}

%----------------------------------------------------------------------------------------

\begin{document}
\begin{newlfm}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

I am Ameen Jaber, PhD candidate,  working with Fadi Zaraket, Assistant professor, 
from the Electrical and Computer Engineering department at the American University of Beirut, 
on Natural Language Processing (NLP) techniques for the Arabic language. 
We are submitting an original manuscript for consideration of publication at 
the IEEE Transactions on Knowledge and Data Engineering journal.

The title of the manuscript is ``MERF: Morphology-based Entity and Relational Entity Extraction Framework for Arabic''.
The manuscript present a set of techniques with a supporting tool to allow user to quickly develop entity and relational
entity extraction application for the Arabic language based on morphology features. 
Existing tools and techniques lack proper support for Arabic morphology analysis which is key for Arabic NLP.
The tool also presents a user friendly interface to help users build their applications.

The paper presents novel specification techniques for Arabic relational entity extraction as it allows 
user defined relations, and code actions based on Arabic morphological analysis.

We highly appreciate your consideration of our manuscript for publication.

%----------------------------------------------------------------------------------------

\end{newlfm}
\end{document}