### 2018-03-14 tex/merf/OldSubmission/Difference/Difference.tex

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm}

\usepackage{charter}

%\newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
%\includegraphics[width=1.2\linewidth]{aub-logo.jpg}}} % Company/institution logo at the top left of the page
\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=50pt} % Slightly decrease the height of the signature field
\newlfmP{addrfromphone} % Print a phone number under the sender's address
\newlfmP{addrfromemail} % Print an email address under the sender's address
\PhrPhone{Phone} % Customize the "Telephone" text
\PhrEmail{Email} % Customize the "E-mail" text

%\lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{Ameen Jaber} % Name

\addrfrom{
\today\\[12pt] % Date
American University of Beirut\\
P.O.Box 11-3875\\
Riad El-Solh / Beirut 1107 2020
}

\phonefrom{(+961)70-962096} % Phone number

\emailfrom{aaj15@aub.edu.lb} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear Jian Pei, editor in cheif, }% Greeting text
\closeline{Sincerely yours,} % Closing text

\begin{document}
\begin{newlfm}

The submitted manuscript 
``MERF: Morphology-based Entity and Relational Entity Extraction Framework for Arabic'', 
significantly extends the work in the attached position 
paper entitled 
``MATAr: Morphology-based Tagger for Arabic'' (AICCSA, 2013). 
MERF is an entity and relational entity extraction framework that 
uses the output of a morphological analyzer to extract entities and 
relations amongst them from text documents. 
MATAr is an Arabic morphology tagger that tags words in text documents 
with their morphological solutions.

In addition, MATAr only supports Boolean formulae defined over morphological features on the word level, 
and MERF extends the support to 
\begin{enumerate}
\item User defined regular expressions over Boolean formulae.
\item User defined relations between subexpressions of the regular expressions. 
\item User defined code actions associated with matches of the subexpressions. 
\item User friendly interfaces to edit the formulae, regular expressions, relations, and actions. 
\item MERF API that allows programmatic access to match morphological features in the code actions. 
\item Automatic cross-reference relation inference that relate entities and relational entities across documents. 
\item User friendly visualizations aids to report the results and facilitate incremental and interactive editing using:
\begin{enumerate}
\item Color and font style sensitive text tagging. 
\item Graph based visualization of morphology and regular expression match trees.
\item Graph based visualization of relational entities and cross reference relations. 
  \end{enumerate}
\end{enumerate}

%----------------------------------------------------------------------------------------

\end{newlfm}
\end{document}