### 2018-03-14 tex/merf/OldSubmission/conclusion.tex

In this work, we presented a morphology-based entity and relational entity information extraction framework for Arabic text; \framework . 
\framework provides a user-friendly interface where the user defines tag types 
and associates them with regular expressions defined over Boolean formulae. 
The Boolean formulae are terms, negations of terms, and disjunctions of terms where terms are matches to Arabic morphological features. 
In addition to the morphological features, \framework introduces $Syn^k$; a semantic feature that relates words based on synonymity. 
The editor allows the user to associate code actions with each regular sub-expression and to define semantic relations between sub-expressions. 
\framework uses an in house Arabic morphological analyzer to compute morphological matches. \framework then computes regular expression matches, 
and then builds the relations. 
We evaluated our work with several case studies comparing the results with existing application-specific techniques. 
The results show good accuracy for \framework compared to manually-coded techniques.

Currently, \framework supports one built in cross-reference predicate 
based on the $Syn^2$ feature. 
In the future, \framework will support user-defined cross-reference predicates. 
Currently the user selects the morphological features to specify the MBF. 
We will explore techniques that can infer the features from example words 
that the user judges as relevant to the basic concept in question.