### 2018-03-14 tex/merf/OldSubmission/discussion.tex

The results show that \framework 
provides a friendly environment to develop entity and relational
entity extraction tasks with acceptable 
accuracy and runtime overheads compared to task specific applications. 
%
\framework requires the user to understand and interact with 
basic linguistic concepts such as readable values of morphological 
features, sequences, repetitions, and bounded repetitions.
The user interacts with the MBF editor to specify basic concepts
and visualize their matches over highlighted text. 
%
Then the user interacts with the MRE editor to specify 
sequences of the concepts and visualize the matches
in a graph form, in conjunction with the highlighted text.

The two levels of interaction allow the user to separate between concepts 
that relate to word features, and more sophisticated entities 
that relate to sequences and context. 
%
The MBF, MRE, and user defined relations 
can be used to generate large annotated corpora in a fast manner. 
\framework visualization can be used later to edit the corpora 
and fix the annotations.

\subsection{ Threats to validity}
As with any GUI tool, it takes time for the user to get acquainted with the 
user friendly interface. 
Relating the subexpressions to their suggested names by \framework might 
not be directly intuitive to the user. 
We address the above two concerns with providing the users with 
a short tutorial video on how to use the tool.