### 2018-03-14 tex/merf/OldSubmission/oldResults.tex

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{\framework~against manually-coded applications: Time, accuracy, and ease of composition}
  \resizebox{\columnwidth}{!}{
    \begin{tabular}{ll|l|l|ll|l}
     \toprule
    \multicolumn{2}{c|}{\multirow{2}{*}{Task}} & \multirow{2}{*}{Build Time} & \multirow{2}{*}{Run Time(s)} & \multicolumn{2}{c|}{Accuracy} & \multirow{2}{*}{Ease of Composition}\\
    \multicolumn{2}{c|}{} & & & Recall & \multicolumn{1}{c|}{Precision} & \\
    \midrule
    \multirow{2}[0]{*}{Narrator Chain} & ANGE & 1-2 month & 1.79 & 1 & 1 & 3000+ lines of code\\
          & framework & 3 hours & 7.24 & 1 & 0.93  & 8 MBFs and 4 MREs\\
    \midrule
    \multirow{2}[0]{*}{Number Normalization} & NUMNORM & 1 week & 0.32 & 0.91 & 0.93  & 500 lines of code\\
          & framework & 1 hour & 1.53 & 0.91 & 0.90  & 3 MBFs/1 MRE/57 lines\\
    \midrule    
    \multirow{2}[0]{*}{Temporal Entity} & ATEEMA & 1-2 month & 2.53 & 0.88  & 0.89  & 1000+ lines of code  \\
          & framework & 3 hours & 3.14 & 0.91  & 0.81  & 3 MBFs and 2 MREs\\
    \bottomrule
    \end{tabular}%
    }
  \label{tab:results}%
\end{table}%