### 2018-04-20 tex/merf/TALMajor/methodology.tex
minor fixes

Figure~\ref{f:overview} illustrates the four processes involved in 
\framework methodology. 
The first process takes Arabic text and provides the user 
with a morphology-based Boolean (MB) formulae GUI. 
The user interactively composes MB-formulae using the GUI and the output 
of the simulator and the $Syn^k$ detector. 
The simulator and the detector apply the formulae over the morphological solutions 
of the Arabic text and produce the MB-formulae tags.

The second process takes the MB-formulae tags and the Arabic text and provides
the user with a morphology-based grammar rule GUI. 
The user interactively composes MB-grammar rules using the GUI and the output 
of the MB-grammar rule simulator. 
The grammar rule simulator applies the rules over the MB-formulae tags and 
produces the MB-grammar rule tags.

The third process takes the MB-grammar rule tags and provides the user with 
a relation and action GUI. 
The user interactively provides (1) the relation definitions and 
(2) the actions in terms of identifiers from the MB-grammar 
rules.
The relation extraction produces the target entities and relational
entities. 
The action execution enriches the entities and the relational entities
with powerful semantics.
For example, users can utilize actions to compute statistical features, 
store intermediate results, 
or apply intelligent entity inference techniques
as we show later in the numerical extraction example of 
Subsection~\ref{subsec:numnorm}.
%
Finally, in the fourth process the user compares the results with golden
reference chunks and visualizes the difference. 
This allows the user to refine the formulae, rules, relations and actions.

After relation extraction, we are interested to relate 
entities that express the same concept.
\framework provides the extended synonym feature of second order as 
a default cross-reference relation ($Syn^2$).
In Figure~\ref{fig:intromotiv}, triggering this feature creates the 
edge labeled with \cci{isSyn} 
between the nodes \cci{Khalifa Tower} and \cci{The building}.

\input{MERFOverview}

The user may inspect and refine the defined formulae, rules and relations and 
the corresponding formulae tags, rule tags, entities and relational entities
either using the GUI or directly through readable output files. 
The files are in the javascript object notation (JSON)~\cite{nolan2014javascript} 
format that is intuitive to read and modify. 
\framework separates the user defined formulae, rules, actions and relations in
a \framework tag type file and the 
matching tags in a tags files. 
The separation serves the user to apply the tag types to multiple case studies 
and to obtain a separate file of resulting tags for each.