### 2018-06-26 tex/sarf/bamainctable.tex

\begin{table}[!tbp]
\begin{minipage}{0.95\textwidth}
\relsize{-1.5} {
  \begin{tabular}{p{2.5cm}ccc}
\hline \hline
   & {\bf Affix } & \textbf{Vocalized} & {\bf Inconsistent tag} \\ 
\hline
\multirow{2}{2.5cm}{ {\bf \ref{inc:plus})}
missing plus in {\bf gloss} tag  of {\bf prefix}}
 & \RL{fl} & \RL{fali} & and/so \fbox{+} for/to + the  \\
% & \RL{fl--l} & \RL{falil} & and/so \fbox{+} for/to + the \\
 & \RL{wbAl} & \RL{wabiAl} & and + with/by \fbox{+} the  \\[6pt]
\multirow{3}{2.5cm}{{\bf \ref{inc:so})} 
  missing alternative {\bf gloss} %for fa
  in {\bf prefix} }
 & \RL{f} & \RL{fa} & and/so  \\ 
 & \RL{fb} & \RL{fabi} & and \fbox{/so} + with/by          \\
 & \RL{fk} & \RL{faka} & and \fbox{/so} + like/such as     \\[6pt]
% & \RL{fl} & \RL{fali} & and \fbox{/so} + for/to           \\
% & \RL{fl} & \RL{fali} & and \fbox{/so} + for/to + the    \\
\multirow{4}{2.5cm}{ {\bf \ref{inc:num_gender_subject}) } 
  gender/number qualifier omitted in {\bf gloss} of subject {\bf suffix} }
 & \RL{n} & \RL{n} & they [fem.pl.] $<$verb$>$                     \\
 & \RL{nhm} & \RL{nahom} & they \fbox{[fem.pl.]} $<$verb$>$ them \\
 & \RL{At} & \RL{At} & [fem.pl.]                                \\
 & \RL{Atk} & \RL{Atka} & \fbox{[fem.pl.]} your                \\[6pt]
% & \multicolumn{4}{c}{48 total inconsistencies} \\ 
\multirow{2}{2.5cm}{
  {\bf \ref{inc:num_gender_object}) } 
  also in {\bf gloss} of object {\bf suffix} }
 & \noVocRL{AnhmA}&  \RL{AnihimA}& them \fbox{(both)}                      \\
 & \noVocRL{tkm}& \utfRL{|Bَتْكُم}& it/they/she $<$verb$>$ you \fbox{(pl.)}  \\[6pt]
% & \multicolumn{4}{c}{39 total inconsistencies} \\ 
\multirow{3}{2.5cm}{
  {\bf \ref{inc:both}) } 
  different ways to express them (both) in {\bf gloss} of {\bf suffix} }
 &\noVocRL{thmA}&  \utfRL{|Bَتْهُما} & it/they/she $<$verb$>$ them (both)    \\
 & \noVocRL{AhmA} &  \utfRL{|BBاهُما} & we $<$verb$>$ 
 \colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}  \\
\vocalize
 & \RL{nAhmA} &  \utfRL{nAhُmA} & we $<$verb$>$ 
\colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}   \\[6pt]
% & \RL{nhmA} &  \utfRL{nَhُmA} & they $<$verb$>$ 
%        \colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}  \\
% & \RL{hmA} &  \utfRL{|BBهُما} & they $<$verb$>$ 
%        \colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}  \\
\multirow{2}{2.5cm}{
  {\bf \ref{inc:dot}) } 
  `.' omitted  after pl in {\bf gloss} }
 & \utfRL{م} & \utfRL{|Bُّم} & you [masc.pl.] $<$verb$>$              \\
% & \utfRL{وا} & \utfRL{|Bُوا} & you [masc.pl\fbox{.}] $<$verb$>$     \\ 
 & \utfRL{ونا} & \utfRL{|Bُونا} & you [masc.pl\fbox{.}] $<$verb$>$ us \\[6pt]
% & \multicolumn{4}{c}{8 total  inconsistencies } \\        
\multirow{2}{2.5cm}{
  {\bf \ref{inc:pos}) } 
  {\bf POS} tag is not same as vocalized 
}
 & \noVocRL{th} & \utfRL{thِ}     & +ti/PVSUFF\_SUBJ:2FS  \\
 &              &                & +\colorbox{shadecolor}{\color{white} hu/}\fbox{hi/}PVSUFF\_DO:3MS \\
% & \utfRL{ونا}  & \utfRL{|Bَوْنا} & +\colorbox{shadecolor}{\color{white} uw/}\fbox{aw/}PVSUFF\_SUBJ:3MP  \\ 
% &              &               & +nA/PVSUFF\_DO:1P \\
%% & \multicolumn{4}{c}{38 total inconsistencies} \\         
\hline \hline
\end{tabular}
}
\end{minipage}
\caption{Sample BAMA inconsistencies}
\label{t:incBAMA}
\end{table}