### 2018-06-26 tex/sarf/cover.tex

\documentclass[11pt,a4paper]{letter}
\usepackage{graphicx}
\usepackage{multirow}

\oddsidemargin -0.25in
\evensidemargin -0.25in
\topmargin -1in
%\topsep 0in
\textheight 10.25in
\textwidth 6.75in
%\parskip 6pt
\longindentation 0.5\textwidth
\parindent 0.4in

\begin{document}

\address{
    \begin{tabular}{p{4.2cm}p{6cm}r}
%    \multirow{3}{*} {
%    \includegraphics[width=4cm]{../AUB-concise-logo-color.jpg} } & &
& & Fadi A. Zaraket \\
& & AUB Faculty of Eng. \& Arch. \\
& & ECE Department \\
%& P.O. Box: 11-0236\\
%& Riad El-Solh, Beirut 1107 2020\\
%& Beirut, Lebanon\\
& & Tel: (+961) 1 350 000, ext: 3484  \\ \hline
& & Email: {\tt fadi.zaraket@aub.edu.lb} \\
\end{tabular}
}

\begin{letter}{}

\signature{Fadi A. Zaraket \\ Assistant Professor
  of Electrical and Computer Engineering}

\opening{Dear concerned at NLE,}

We appreciate your consideration of our submission entitled {\em Sarf: Application Customizable Efficient Arabic Morphological Analyzer}. 
The paper describes an Arabic morphological analyzer that has been used effectively and efficiently in several Arabic NLP 
applications for information extraction from Arabic text. 
The paper describes the contributions of Sarf as compared with existing work. 
Sarf provides application customizable analysis through an API that allows the NLP application developer to control and refine the 
analysis on the fly. 
Sarf refines the lexicons of existing analyzers, reduces the required size and maintenance by using agglutinative and fusional concatenation rules and solves morphological feature inconsistencies in the existing analyzers. 
Sarf also solves the token segmentation problem which is a hindrance to machine learning with current analyzers.

We look forward for NLE to consider our submission as Sarf bridges the gap between computational linguistics and practical applications, 
provides an efficient engineering implementation for the Arabic morphological analysis problem, and is similar to 
morphological analyzers and lexical resources for other semitic languages such as Hebrew that NLE published before.

\closing{Best regards,
    \\~
}
\vspace{-2in}

\end{letter}

\end{document}

\end{document}