### 2018-06-26 tex/sarf/placestable.tex

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}
  \begin{minipage}{\textwidth}
%  \relsize{-2} {
    \begin{tabular}{l|ccc|ccc|ccc}
    \hline \hline
    \multicolumn{1}{c}{} & \multicolumn{3}{c}{Kafi} & \multicolumn{3}{c}{Istibsar} & \multicolumn{3}{c}{Ibn Hanbal} \\
    \hline
          & correct & detected & required & correct & detected & required & correct & detected & required \\
    Places & 20    & 213   & 32    & 26    & 113   & 29    & 11    & 127   & 44 \\
    Possessive & 6     & 21    & 7     & 15    & 20    & 16    & 3     & 18    & 8 \\
    vicinity of name & 8     & 180   & 13    & 18    & 99    & 22    & 9     & 111   & 20 \\
    \hline \hline
    \end{tabular}%
 %   }
    \end{minipage}
  \caption{Sarf results of extracting places, places with possessive suffix, and places in vicinity of proper names.}
  \label{tab:placesRes}%
\end{table}%