### 2018-06-26 tex/sarf/rev1.tex

\documentclass[11pt]{article}
\usepackage[letterpaper,margin=1.5in,left=35mm,right=20mm, top=20mm, bottom=20mm ]{geometry}

\usepackage{url}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage{amssymb}

% used for cross reference from psq2aig.tex
\usepackage{xr} 
\externaldocument{psqBMCaig}
% You ou can also add a prefix: 
%\externaldocument[I-]{chapterI}
% then you can reference it as \ref{I-a}

\RequirePackage{xspace}
\linespread{1.5}

\newcommand{\cci}[1]{\mbox{\textup{\small \texttt{#1}}}}
\newcommand{\psqlanguage}{\ensuremath{\mathit{J}}\xspace}
\newcommand{\psqcore}{\ensuremath{\mathit{Jcore}}\xspace}
\newcommand{\psqtool}{\ensuremath{\{\cal P\}{\cal S}\{\cal Q\}}\xspace}
\newcommand{\oneloop}{\ensuremath{\mathit{\cal O}lp}\xspace}
\newcommand{\true}{\ensuremath{\mathit{true}}\xspace}
\newcommand{\false}{\ensuremath{\mathit{false}}\xspace}
\newcommand{\itodo}[1]{\todo[inline,color=green]{#1}}
\newcommand{\Pm}{\ensuremath{\mathcal{S}}\xspace}         % program

\title{Sarf: Fast and Application Customizable Arabic Morphological Analyzer
 \\
~~ \\
  Revision summary }
\date{}

\begin{document}
\maketitle

{\noindent Dear Editor-in-Chief and Expert Reviewers,}

We would like to thank you for your recommendations, valuable comments, 
and constructive observations. 
In this revised version of the paper we made every effort to address the 
concerns brought up in your review. 
This document includes a summary of our changes and lists each of the 
reviewers’ comments followed by a description of how it was addressed.

For the convenience of the reviewers, we also provide online 
(\url{http://research-fadi.aub.edu.lb/psqaigfiles/psqaig-tse-revised-annotated.pdf})  
a version of the revised paper with annotations that point to changes and 
revisions and relate them to the concerns of the reviewers.

Thank you,
\newline
\hrule

\input{editorComments.tex}
~\newline
\hrule

\input{Reviewer1.comments}
~\newline
\hrule

\input{Reviewer2.comments}
~\newline
\hrule

\input{Reviewer3.comments}
~\newline
\hrule

\newsavebox\mytempbib
\savebox\mytempbib{\parbox{\textwidth}{
%\input{psqBMCaig.bbl}
}}

\end{document}