### 2018-06-26 tex/sarf/samainctable.tex

\begin{table}[!tbp]
\resizebox{0.95\textwidth}{!}{
%\begin{minipage}{0.8\textwidth}
\relsize{-1.5} {
\begin{tabular}{p{2.5cm}cccc}
\hline \hline
 & {\bf Affix } & \textbf{Vocalized} & {\bf Inconsistent tag} & \\ \hline
\multirow{2}{2.5cm}{
  {\bf \ref{inc:relaxed})} 
  missing standalone alef with no hamza {\bf prefix}  forms
}
% & \RL{A} & \utfRL{أَ} & I \\ 
 & \RL{A} & \utfRL{آ} & I  \\ 
% & \RL{A} & \utfRL{أُ} & I  \\
 & \RL{sA} & \utfRL{سَأُ} & I \\[6pt]
\multirow{2}{2.5cm}{
  {\bf \ref{inc:by}) } 
   additional by in {\bf gloss} 
}
 & \RL{w} & \RL{wa} & with      \\ 
 & \RL{wAl-} & \RL{waAl-} & with\colorbox{shadecolor}{\color{white} /by} + the \\[6pt]
\multirow{1}{2.5cm}{ {\bf \ref{inc:fa}) } 
  additional space in vocalized form 
}
& \RL{f-} & \colorbox{shadecolor}{~}\RL{fa-}\colorbox{shadecolor}{~} \\[6pt]
\multirow{1}{2.5cm}{ {\bf \ref{inc:wafa})} 
   wrong {\bf prefix}
}
& \RL{wf-}& \RL{wafa-} & and + so/and \\[6pt]
\multirow{3}{2.5cm}{ {\bf \ref{inc:def}) } 
  missing definite indicator in {\bf suffix gloss} 
}
 & \utfRL{آت} & \utfRL{آتِ} & [fem.pl.] + [def.acc.]                               \\
 & \utfRL{اتك} & \utfRL{اتِكِ} & [fem.pl.] + [def.acc.] + your [fem.sg.]            \\
 & \utfRL{آتك} & \utfRL{  آتِكِ} & [fem.pl.] + [\fbox{def.}acc.] + your [fem.sg.]  \\[6pt]
% & \multicolumn{4}{c}{33 other inconsistencies} \\ 
\multirow{2}{2.5cm}{ {\bf \ref{inc:num_gender}) }
  omitted gender/num  qualifier in {\bf gloss}
}
 & \utfRL{اك}&  \utfRL{|BBاكِ}& we [verb] + you \fbox{[fem.sg.]}          \\
 & \utfRL{نهم}&  \utfRL{|Bْنَهُم}& they \fbox{[fem.pl.]} [verb] + them      \\[6pt]
% & \multicolumn{4}{c}{72 total inconsistencies} \\ 
\multirow{3}{2.5cm }{  {\bf \ref{inc:both:sama}) } 
   different ways to express them (both) in {\bf gloss} of {\bf suffix} 
}
 & \noVocRL{thmA}&  \utfRL{|Bَتْهُما} & it/they/she [verb] them (both)   \\
 & \noVocRL{AhmA} &  \noTrutfRL{|BBاهُما} \noArutfRL{ّاهُما}& we [verb] 
                \colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}   \\
 & \noVocRL{nAhmA} &  \noVocRL{nAhomA} & we [verb] 
                \colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}   \\[6pt]
% & \noVocRL{nhmA} &  \noVocRL{nahomA} & they [verb]
%                \colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}  \\
% & \noVocRL{hmA} &  \utfRL{|BBهُما} & they [verb[ 
%\colorbox{shadecolor}{\color{white} (both of)} them \fbox{(both)}  \\
\multirow{3}{2.5cm}{ {\bf \ref{inc:num_gender_differences}) } 
  leftover BAMA style tags  in {\bf gloss}
}
 & \utfRL{كم} & \utfRL{كُم} & he/it [verb] + you [masc.pl.]        \\
 & \utfRL{تكم} & \utfRL{تُكُم} & I [verb] + you 
                \colorbox{shadecolor}{\color{white}(pl.)} \fbox{[masc.pl.]}  \\
 & \utfRL{كن} & \utfRL{|Bُّكُنَّ}  & I [verb] + you 
                \colorbox{shadecolor}{\color{white}(women) } \fbox{[fem.pl.]}   \\[6pt]
% & \multicolumn{4}{c}{13 other inconsistencies } \\                
\multirow{3}{2.5cm}{ {\bf \ref{inc:jus}) } 
  indicative {\bf gloss} with jussive {\bf POS}
}
% & ~ & ~ & [jus.] ~ ~ ~ (null)/IVSUFF\_MOOD:J  \\ 
 & \utfRL{نا} & \utfRL{نا} & 
       IVSUFF\_MOOD:J,
       \colorbox{shadecolor}{\color{white}[ind.]} \fbox{[jus.]} + us  \\
 & \utfRL{ني} & \utfRL{نِي} & 
       IVSUFF\_MOOD:J,
       \colorbox{shadecolor}{\color{white}[ind.]} \fbox{[jus.]} + me \\[6pt]
\multirow{3}{2.5cm}{ {\bf \ref{inc:dot:SAMA}) } 
   omitted `.'  in {\bf gloss} 
}
 & & \utfRL{|Bُ} & [def.nom.] & \\ 
 & \utfRL{ه} & \utfRL{|Bُهُ} & [def.nom \fbox{.}]  \\
 & \utfRL{تكن} & \utfRL{|Bَتُكُنَّ} & 
[fem.sg.] + [def.nom \fbox{.}] + your [fem.pl \fbox{.}]  \\[6pt]
% & \multicolumn{4}{c}{36 other inconsistencies } \\
\vocalize
\multirow{2}{2.5cm} { {\bf \ref{inc:pos:SAMA})} 
  shadda inconsistent in {\bf POS} 
}
 & \utfRL{ي} & \utfRL{يَ} & ya/POSS\_PRON\_1S \\
 & \utfRL{ي} & \utfRL{|Bِyya} & 
 % iy/NSUFF\_MASC\_PL\_NOM  +\colorbox{shadecolor}{\color{white}$\sim$a/}\fbox{ya/}POSS\_PRON\_1S  \\
 $\ldots$ +\colorbox{shadecolor}{\color{white}$\sim$a/}\fbox{ya/}POSS\_PRON\_1S  \\
\hline \hline
\end{tabular}
}
}
%\end{minipage}
\caption{Sample SAMA inconsistencies}
\vspace{-2em}
\label{t:incSAMA}
\end{table}