### 2018-06-26 tex/sarf/sarf.tex

In this Section, we describe the methodology used to construct the affix and stem data structures. 
We also discuss the advantage of choosing these data structures over the ones used in the literature. 
We also describe the construction of the system of Sarf by combining the different data structures.

\subsection{Affix and stem data structures}
%\label{subsec:dfsas}

We analyzed the prefixes and suffixes of 
Buckwalter~\shortcite{Buckwalter:02}
and constructed the agglutinative affixes
into directed acyclic graphs (DAG) similar to 
those in Figures~\ref{f:example}(a) and~\ref{f:example}(c).
The use of the DAG data structure provides compactness as well as 
linear time traversal.
This is computationally superior to the 
approach of Buckwalter where the analyzer considers
all possible substrings and looks them up in the affix tables.

The affix structures encode $L_{pp}$ and $L_{xx}$. 
An accept node in the DAG corresponds to the last letter 
of an affix in the affix lists. 
The accept nodes produce as output the POS and other tags
associated with the node. 
Consider the affix \noTrutfRL{وسيـ} of the word \RL{wsyl`bhA}. 
Possible accept nodes for this input in ${\cal P}$ are for 
\noTrutfRL{و} and \noTrutfRL{يـ}. 
However, only the accept node of \noTrutfRL{يـ} will yield a valid 
solution for the input word \RL{wsyl`bhA}.

%\subsection{Stem transducer}
%\label{subsec:stemFSA}

We built our stem lexicon using the stem lexicon of 
Buckwalter. 
Sarf extends the lexicon of Buckwalter~\shortcite{Buckwalter:02} with
proper names and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.
%We augmented the lexicon with a set of proper names and
%a set of location names which we 
%obtained from online and biblical sources.

The stems share lots of sub-strings. We encoded them in
an efficient double array trie structure~\citep{Aoe:89}. 
The accept nodes are the terminal nodes corresponding to the last 
letters in stems. 
This approach is superior to Buckwalter since it consists of
a linear walk in the trie while with Buckwalter we need 
a number of hash lookups in the order of all possible partitions
of the string. 
For example given the word \RL{'kl}, our approach requires three 
steps to check that this is a valid stem. 
However, buckwalter needs to partition this word into six different partitions 
where each partition contains possible prefix, stem, and suffix.

\subsection{Combined system}
%\label{subsec:combined}

\begin{figure}
\begin{minipage}{0.6\textwidth}
\begin{Verbatim}[fontsize=\relsize{-2},
frame=topline,framesep=4mm,label=\fbox{NDSarf algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
NDSarf(string $L$, Interface $A$) 
  TraversalVec $M$; -- collection of traversals
  Traversal $\Phi_1$;
  $M$.insert($\Phi_1$);
  foreach $c$ in $L$ \{
    foreach $\Phi$ in $M$ \{
      if ( $\Phi$.node.isAccept() ) \{
        if ($\Phi$.isSuffix())
          if ($c$.isWhite()) or 
            $A$.report();
          if ($\Phi$.lastChar().isNonConnecting())
            $A$.report();
        $\Psi$ = $\Phi$.clone();
        $M$.insert($\Psi$); \}

      if ( $\Phi$.isWalkable($c$) ) \{
        $\Phi$.child($c$);
      \} else \{
        $M$.remove($\Phi$);
        $\Phi$.die();
      \} \} 
    $A$.control($M$, $c$); \}
\end{Verbatim}
\end{minipage}
\caption{Sarf algorithm}
\label{f:ndsarf}
\end{figure}

The algorithm \CodeIn{NDSarf} shown in Figure~\ref{f:ndsarf}
takes as input a text string $L$ and a user-implemented interface
$A$. 
It starts with one traversal $\Phi_1$ and inserts it into a collection $M$. 
The algorithm reads a letter $c$ at a time from $L$
and iterates over all the traversals in $M$. 
For each $\Phi$ in $M$, 
if $\Phi$ is in an accept node and it is in the suffix
phase, then the algorithm checks whether it should report
a full word. 
This happens when $c$ is a white space, a delimiter, 
or when the preceding letter %leading to the current state
was a non-connecting letter.

In all cases, if $\Phi$ is in an accept node, 
it spawns another traversal $\Psi$ and adds it to $M$. 
Notice that spawning a new traversal requires only keeping
the current node of the traversal since all traversals share the
structures of the three transducers.
If $c$ leads to a reject, then $\Phi$ dies 
and we remove it from $M$. 
Otherwise, $\Phi$ proceeds using the $c$ edge.
After iterating over all traversals in $M$, we invoke the 
user-implemented interface $A$
and grant it full access and control over
all traversals in $M$.