### 2018-06-26 tex/sarf/solutiontable.tex

\begin{table}[tb]
\centering
\begin{minipage}{.9\textwidth}
\caption{\label{t:solexample}Morphological solution for \utfRL{وسيلعبونها}(and they will play it)}
\begin{tabular}{lccc} \hline
  & suffix  \noTrutfRL{ونها} & stem  \noTrutfRL{لعبـ}  & prefix \noTrutfRL{وسيـ} \\ \hline
       POS & \cci{IVSUFF\_SUBJ:MP\_MOOD:I+IVSUFF\_DO:3FS} & \cci{VERB\_IMPERFECT} & \cci{CONJ+FUT+IV3MP}  \\
  Transliteration & \cci{uwnahA} & \cci{loEab} & \cci{wa+sa+ya}  
  \\
           Gloss  & \cci{[MASC.PL.]+it/them/her}.  & \cci{play} & \cci{and they will} \\
  \hline% \hline
\end{tabular}
\end{minipage}
\end{table}