### 2018-06-27 tex/merf/TALMajor/abstract.tex
french translation and shrink first page

Rule-based techniques to extract relational entities from 
documents allow users to specify desired entities with 
natural language questions, finite state automata, regular expressions and
structured query language. 
They require linguistic and programming expertise and 
lack support for Arabic morphological analysis.
We present a morphology-based entity and relational entity extraction 
framework for Arabic (\framework).
\framework requires
basic knowledge of linguistic features and regular expressions, 
and provides the ability to interactively specify
Arabic morphological and synonymity features, 
tag types associated with regular expressions, and
relations and code actions defined over matches of subexpressions. 
\framework constructs entities and relational entities from matches 
of the specifications.
We evaluated \framework with several case studies.
The results show that \framework requires shorter development time and effort compared to existing application specific techniques and produces 
reasonably accurate results within a reasonable overhead in run time.